### 2021-06-30 tables/accuracy_table.Rmd
Rename folder

```{r setup, include = FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
source(here("code/03_build_models.R"))
```
```{r make_table, include = FALSE} 
# Load packages and functions ---------------------------------------------

library(here)
library(glue)
library(kableExtra)
source(here("code/functions/table_functions.R"))

# Load models data --------------------------------------------------------

source(here("code/03_build_models.R"))

# Build table -------------------------------------------------------------

# Create a list with equation coefficients and accuracy indices
placements <- c("Ankle", "Lower back", "Hip")
coeffs <- list(
  ankle = list(
    resultant = GRF_coefficients_jumping$ankle_resultant,
    vertical = GRF_coefficients_jumping$ankle_vertical
  ),
  back = list(
    resultant = GRF_coefficients_jumping$lower_back_resultant,
    vertical = GRF_coefficients_jumping$lower_back_vertical
  ),
  hip = list(
    resultant = GRF_coefficients_jumping$hip_resultant,
    vertical = GRF_coefficients_jumping$hip_vertical
  )
)
accuracy <- list(
  ankle = list(
    resultant = GRF_accuracy_jumping$ankle_resultant,
    vertical = GRF_accuracy_jumping$ankle_vertical
  ),
  back = list(
    resultant = GRF_accuracy_jumping$lower_back_resultant,
    vertical = GRF_accuracy_jumping$lower_back_vertical
  ),
  hip = list(
    resultant = GRF_accuracy_jumping$hip_resultant,
    vertical = GRF_accuracy_jumping$hip_vertical
  )
)

a <- c("Ankle", "Lower back", "Hip")
b <- c("ankle", "back", "hip")
table <- map2_dfr(a, b, ~ accuracy_table(.x, coeffs[[.y]], accuracy[[.y]]))
```
```{r table, echo = FALSE}
kbl(table, booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```