### 2022-03-04 manuscript/manuscript.Rmd
Revise the manuscript

---
title: "Using raw accelerometer data to predict high-impact mechanical loading"
output:
  bookdown::pdf_document2:
    toc: false
    includes:
      in_header: "preamble.tex"
    latex_engine: lualatex
  word_document:
    reference_docx: style_reference.docx
always_allow_html: true
bibliography: bibliography.bib
csl: nlm.csl
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, results = "hide", message = FALSE, warning = FALSE
)
```

```{r source}
library(here)
library(bookdown)
library(broman)
source(here("code", "02_explore.R"))
```

```{r GRF_accuracy}
load(here("output", "prediction_models.rda"))
res_GRF_accuracy <- purrr::map(
  accuracy_res_GRF_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "resultant"
  )
ver_GRF_accuracy <- purrr::map(
  accuracy_ver_GRF_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "vertical"
  )
GRF_accuracy <- rbind(res_GRF_accuracy, ver_GRF_accuracy)
hip_GRF_accuracy <- GRF_accuracy %>%
  filter(placement == "hip" & vector == "resultant")
```

```{r LR_accuracy}
load(here("output", "prediction_models.rda"))
res_LR_accuracy <- purrr::map(
  accuracy_res_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "resultant"
  )
ver_LR_accuracy <- purrr::map(
  accuracy_ver_LR_models,
  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
) %>%
  purrr::map_dfr(rbind) %>%
  dplyr::mutate(
    placement = c("ankle", "lower back", "hip"),
    vector = "vertical"
  )
LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
hip_LR_accuracy <- LR_accuracy %>%
  filter(placement == "hip" & vector == "resultant")
```

```{r ba_agreement}
load(here("output", "loocv_data.rda"))
percent_inside_loa <- function(data) {
  lower <- lvmisc::loa(
    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
  )$lower
  upper <- lvmisc::loa(
    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
  )$upper
  diff <- dplyr::transmute(
    data, diff = .data[[".actual"]] - .data[[".predicted"]]
  )
  diff <- dplyr::mutate(
    diff,
    inside = as.factor(ifelse(diff > lower & diff < upper, "yes", "no"))
  )
  round(prop.table(table(diff$inside))[[2]] * 100, 0)
}

ba_agreement <- min(
  map_dbl(cv_res_GRF_models, percent_inside_loa) %>% min(),
  map_dbl(cv_ver_GRF_models, percent_inside_loa) %>% min(),
  map_dbl(cv_res_LR_models, percent_inside_loa) %>% min(),
  map_dbl(cv_ver_LR_models, percent_inside_loa) %>% min()
)
```

```{r ba_regression}
source(here("code", "04_analyse_results.R"))
ba_regression_ankle_R2 <- c(
  ba_regression_res_GRF$ankle["adj.r.squared"],
  ba_regression_ver_GRF$ankle["adj.r.squared"],
  ba_regression_res_LR$ankle["adj.r.squared"],
  ba_regression_ver_LR$ankle["adj.r.squared"]
) %>%
  as_vector()
```

<!--
<center>
Lucas Veras$^{1,2}$, Florêncio Diniz-Sousa$^{1,2}$, Giorjines Boppre$^{1,2}$, Vítor Devezas$^3$, Hugo Santos-Sousa$^3$, John Preto$^3$, João Paulo Vilas-Boas$^{4,5}$, Leandro Machado$^{4,5}$, José Oliveira$^{1,2}$, Hélder Fonseca$^{1,2}$
</center>

\bigskip

1 - Research Center in Physical Activity, Health and Leisure (CIAFEL), Faculty of Sport, University of Porto, Porto, Portugal

2 - Laboratory for Integrative and Translational Research in Population Health (ITR), University of Porto, Porto, Portugal

3 - General Surgery Department, São João Medical Center, Porto, Portugal

4 - Center of Research, Education, Innovation and Intervention in Sport (CIFI2D), Faculty of Sport, University of Porto, Porto, Portugal

5 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal

\leftskip=0.5in \rightskip=0.5in

\bigskip
-->

# Introduction {.unnumbered}

Bone responds to mechanical loading by altering its properties, such as density, geometry, and strength [@Turner2003; @Fonseca2014; @Santos2017]. Physical activity is a way of eliciting mechanical stimuli, and it is well established to improve bone health. However, there is no consensus in which characteristics of the physical activity-related mechanical stimulus are better associated with bone health benefits. Nevertheless, there is evidence that high-impact exercise, although performed scarcely throughout the day, may play an important role in inducing positive bone adaptations [@Vainionpaa2006; @Stiles2017].

But, even when knowing precisely which the best loading regimen is to improve bone health, the question of how to measure and monitor this load remains. Bone mechanical loading is usually measured in laboratory settings, using instruments such as force plates (FP) [@Medved2000]. Measurements in these instruments, that are confined to a laboratory space, are not applicable in controlling the loads that a person is routinely subjected to in free-living contexts. To surpass this issue, in the last few years, methods using wearable devices such as accelerometers have been proposed to measure mechanical loading, more precisely,  ground reaction force (GRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020] and loading rate (LR) [@Fortune2014; @Veras2020]. These devices are light, small, can be used throughout days, and have a relatively affordable price. These characteristics make the accelerometers well-suited instruments to measure mechanical loading on a daily basis.

However, the validity of the accelerometer-predicted mechanical loading values depends on some factors, as the sample characteristics [@Welk2005]. In fact, due to this, several calibration studies have been done in several populations, such as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020]. Another important factor that influences the validity of the predictions is the type of activities performed in the calibration study [@Bassett2012]. Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed activities such as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014]. Nevertheless, activities eliciting higher impact loads have not yet been tested.

Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activities in subjects from normal weight to severe obesity.

# Methods {.unnumbered}

A total of `r max(sample_size$pGRF_N)` adults were enrolled in the study (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD). The protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent. No participant reported any musculoskeletal limitations. The measurement of height (stadiometer model 213; Seca, Hamburg, Germany) and body mass (digital scale model 899; Seca, Hamburg, Germany) was accomplished according to standard practice.

In this study, jumps were chosen as the high-impact activities to be performed. After familiarization, participants completed a series of trials involving jumps of different types and heights. First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments. Then, box jumps were performed from the floor to 5cm, 15cm, and 30cm high boxes. Finally, continuous jumps were carried out at two different heights: 5cm and 15cm. Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height. The landing phase of the drop jumps, the take-off phase of the box jumps, and both phases of the continuous jumps were recorded in FC (1000Hz; AMTI Corporation, Watertown, USA). Participants were instructed to jump with no restrictions to arms movement, land two-footed, and then remain stationary on the force plate for five seconds (except in the continuous jumps).

Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus). Monitors worn at hip and lower back were positioned on the same elastic belt with clips, and the ankle-worn monitor was fixed by an elastic belt and adhesive tape. Activity monitors were always positioned to ensure the alignment of the accelerometer vertical axis and the standing body craniocaudal axis.

The activity monitors used include a primary and a secondary triaxial accelerometers. As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, only data from this accelerometer were used in this study, to improve data processing replicability. Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors. GRF was expressed in N and acceleration (ACC) in gravitational ACC units (1*g* = 9.807 m·s^-2^).

The MATLAB software (Version 2019a, Mathworks, Natick, USA) was used to process data from both FP and accelerometers. Data were processed according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz). Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude. This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT). Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$). After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary. Next, peak ACC (pACC) was determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps. Peak GRF (pGRF) was then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type. The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1). The peak LR (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).

\[f_i^{'} = \frac{f_{i + 1} - f_{i - 1}}{t_{i + 1} - t_{i - 1}}\]

\[f_{i_{max}}^{'} = max(f_i^{'})\]

Finally, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each jump type and height for each accelerometer placement were extracted and used in all remaining analyses.

Statistical analyses were performed through the R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria). All code used to generate the results can be found in an open platform (https://github.com/verasls/jumping_GRF_ACC)(ref). Statistical significance was set as $\alpha$ = 0.05.

Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back, and  hip). Body mass was tested as a predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models. These predictors were entered as fixed effects and kept in the model, as all have shown to be significant (*p* < 0.05). Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models. Final models were chosen according to the -2 log-likelihood statistic [@Field2012]. The conditional coefficient of determination (R^2^), which estimates the variance explained by both fixed and random factors [@Nakagawa2013], was computed.

Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012]. This method consists in separating data from one of the subjects into a testing dataset, and keeping the remainder in the training dataset. Then, a new model is developed using the training dataset, with the same parameters that were defined using the whole sample. This new model is then applied to predict the outcome for the subject in the testing dataset. The whole process is repeated once for each participant (`r max(sample_size$pGRF_N)` times). Data from the testing dataset were used in all subsequent analyses.

The quality of the models was then evaluated both numerically and graphically. The numerical assessment was made by computing the following accuracy indices: the mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE). Despite the absence of a standard criterion for establishing an acceptable error in mechanical loading prediction, we regarded a MAPE <8\% to be an accurate prediction based on past research [@Fortune2014; @Neugebauer2014; @Ancillao2018, @Veras2020]. To visualize the agreement between the values predicted by the models and measured by the FP, Bland-Altman plots were drawn. This type of plot is made by plotting the mean of the actual and predicted values on the x axis and their difference on the y axis. Then, the bias was calculated as the mean of these differences and plotted as a continuous horizontal line, and the limits of agreement were defined by bias ± 1.96 standard deviation and plotted as horizontal dashed lines [@Bland1986]. To further test if there were any consistent under- or over-estimation of the values by the models, one-sample *t*-tests were done to check whether the bias were different than zero. Finally, linear regressions were employed to examine if the difference between actual and predicted values were affected by their magnitude, verifying, therefore, whether the prediction was constant throughout the magnitude range that was tested [@Giavarina2015].

Lastly, our pGRF and pLR prediction models were compared with a previously published equation by our group [@Veras2020]. As, for the best of our knowledge, there is no mechanical loading prediction equation developed for jumping activities, we chose to compare it with previous models from our group that were validated in a similar sample. The models used in the comparison were the ones predicting pGRF and pLR from both the resultant and vertical vectors of accelerometers worn at lower back and hip. The MAE, MAPE and RMSE were calculated using the values predicted using Veras et al. [@Veras2020] equation.

# Results {.unnumbered}

\begin{figure}[ht]
	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
	\centering
	\begin{measuredfigure}
		\includegraphics[width=\textwidth]{fig1.png}
    \caption{Distribution of the resultant vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination. Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
	\end{measuredfigure}
\end{figure}

In our protocol, three distinct jump types were performed, in several different heights. This variety of jump trials produced a varying response in both the FP- and accelerometer-measured variables. These responses in pGRF, pLR, pACC, and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector). Panels A and B show the FP-measured variables normalized by the subjects body mass, the pGRF and pLR, respectively. It can be seen that in the drop jump trials, there is a tendency to increase the loading magnitude the higher the jump, and in the box jumps this tendency is lower. As for the continuous jumps, both assayed heights elicit reasonably constant loadings. The same pattern can be observed in the accelerometer-measured variables, pACC and pAR, in Panels C and D, respectively. Also, regarding the influence of the accelerometer placement, in most of the trials accelerometers worn at hip and lower back produced approximately similar values, with a tendency to higher dispersion in data derived from lower back accelerometers.

```{r accuracy_table, results = "show"}
source(here("tables", "tab1.R"))
tab1
```

Accelerometer-based mechanical loading prediction equations were developed based on linear mixed models. The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back, and hip, and from the resultant and vertical vectors are presented in Table 1. The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, showing that the models can explain at least `r round(min(GRF_accuracy$R2_cond), 1) * 100`\% of the pGRF variance. The observed MAPE values were between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which falls outside of the proposed threshold of 8\%. Nevertheless, these MAPE values are similar to the results usually found in accelerometer-based energy expenditure prediction [@Diniz-Sousa2020], which have widespread use in research. Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%), and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N). Compared to the pGRF models, the pLR equations yielded poorer results, which is expected based on past research [@Fortune2014; @Veras2020]. Their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)` and MAPE from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%. However, similar to the pGRF models, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector of hip-worn accelerometers, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\%, and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.

\begin{figure}[ht]
	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
	\centering
	\begin{measuredfigure}
		\includegraphics[width=\textwidth]{fig2.png}
		\caption{Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers. Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation). Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.}
	\end{measuredfigure}
\end{figure}

To visualize the agreement between the pGRF and pLR values measured by the FP and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vertical vectors and for all accelerometer placements. Figure 2 shows the Bland-Altman plots for the pGRF (Panel A) and pLR (Panel B) resultant vector of hip-worn accelerometers. Both plots have shown good agreement levels, as at least `r ba_agreement`\% of the data points are within the limits of agreement and data tended to cluster around zero. Bland-Altman plots for the other accelerometer placements and vectors have shown the same patterns described previously, and can be found in the Supplemental Figure S2. Also, as could be observed by the one-sample *t*-tests results, none of the equations presented a bias different than zero (*p* > 0.05). Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influences the prediction error. However, as the linear regressions presented a very small R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.

```{r accuracy_table2, results = "show"}
source(here("tables", "tab2.R"))
tab2
```

\begin{figure}[ht]
	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
	\centering
	\begin{measuredfigure}
		\includegraphics[width=\textwidth]{fig3.png}
    \caption{A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points). Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
	\end{measuredfigure}
\end{figure}

Although the accuracy of our equations showed to be beyond the established 8\% threshold, when applying a model not developed for jumping activities, we can see that the accuracy tremendously decreases, reaching values of MAPE around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[1:4]) / 50) * 50`\% for pGRF prediction and around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[5:8]) / 10) * 10`\% for pLR prediction as can be observed in Table 2. Furthermore, applying to our accelerometer-recorded jumping data a model that is not specific for this activity yields unreasonable results, such as the negative values for pGRF that can be observed in Figure 3, Panel A.

Finally, it is important to mention that an R package was developed [@Veras2021] to make it easier to apply all the necessary data processing steps to use the prediction equations. The package website (https://lveras.com/impactr) has all the installation instructions and documentation necessary to apply this method on your own data.

# Discussion {.unnumbered}

This study aimed to develop accelerometer-based mechanical loading prediction equations for jumps of diverse types and heights, testing different accelerometer placements and using both the resultant and vertical vectors. The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activities, such as walking and running. Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for lower impact activities in a sample similar to the current study yielded far worse results in terms of prediction accuracy. Also, as expected, the pLR prediction models performed poorly compared to the pGRF equations.

Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly at higher magnitudes [@Vainionpaa2006; @Stiles2017]. This type of high-intensity impact activities have shown to be more associated with bone health improvements, even if they are usually performed in low quantities throughout the day. In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density. Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps [@Vainionpaa2006; @Rowlands2012]. Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activities, such as walking and running, that elicit relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].

The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed [@Welk2005]. Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities. Even though the cross-validation accuracy of such a model was poorer compared with other models developed for activities such as walking and running, as demonstrated by our results, the use of these models is crucial to achieve an acceptable prediction accuracy. It was demonstrated by testing a model previously developed by our group [@Veras2020] in a similar sample, but on walking, on the accelerometer-recorded jumping data from the current study. The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force. These negative pGRF values were mainly as a result of a predictor, the second-degree polynomial of the acceleration, possessing a negative coefficient. While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs. pACC relationship during jumping, and, therefore, only impairs the accuracy of these equations when applyed to this kind of activity.

Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beginning of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020]. This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequencies, that are usually 100Hz at maximum. Based on this and on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations. Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].

In our protocol, three accelerometer placements (ankle, lower back, and hip) and two vectors (resultant and vertical) have been tested. Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be drawn. First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a slightly better accuracy compared the other placements tested. This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions. Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so. Second, as the magnitude of the resultant vector of both pGRF and pLR, in the activities that were tested, is derived mostly from the vertical vector, and there were no considerable differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading. Besides that, while models using data from the vertical vector allow the use of uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct accelerometer axes orientation, which cannot be guaranteed in free-living data collection [@Welk2005].

The development of accelerometry-based mechanical loading prediction models was done with the direct purpose of allowing to assess the loading beared by the skeletal system during daily physical activity. This would allow some clinical applications. For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters. This would allow a more complete understanding on how mechanical loading influences bone health, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits. Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability [@Veras2021]. This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.

Although the models presented in this study showed to be valid, some limitations ought to be considered. First, the choice of activity to elicit high-impact loading was jumping, and several combinations of jump type and height were tested. Nevertheless, the results presented here cannot be assured for other high-impact activities, or even for other types of jumps that were not tested. Second, no external sample for model validation was recruited. However, the leave-one-out cross-validation method that was used is the recommended in these situations [@Staudenmayer2012]. Third, to apply these models in accelerometer-recorded free-living physical activity data, there is a need to identify, in the acceleration signal, the periods in which the activity being performed is jumping. Therefore, methods for physical activity pattern detection should be used in combination with these models.

Mechanical loading prediction equations based on accelerometer data have already been proposed for several populations [@Neugebauer2012; @Neugebauer2018; @Veras2020] and using an accelerometer in different body placements [@Fortune2014; @Veras2020]. What these past few studies have in common is that they validated their models using locomotion activities, such as walking only or walking and running. The current study, by setting means to predict mechanical loading in several types and heights of jumps, broadens the application of such models, getting a step closer to the ability to accurately identify daily-living mechanical loading due to physical activity. This ability would allow the monitoring of loading in different contexts and to more fully understand the relationship between mechanical stimulation and its consequent effect on bone health.

\newpage

# References {.unnumbered}

<div id="refs"></div>

\newpage

# Figure captions {.unnumbered}

**Figure 1:** Distribution of the resultant vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination. Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.

**Figure 2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers. Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation). Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.

**Figure 3:** A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points). Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.

**Supplemental Figure S1:** Distribution of the vertical vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination. Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.

**Supplemental Figure S2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F). Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ±1.96 standard deviation). Abbreviations: pGRF, peak ground reaction force.